---
title: "Homework 1"
author: "Aakriti Gupta"
date: 2018-09-21
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library (knitr)
library(dplyr)
```

## **Problem 1**
We will first create a dataframe containing variables of four different types: 
(1) random sample of 10 numbers from a uniform distribution between 0 to 5
(2) logical vector indicating whether the numbers from the sample are greater than 2
(3) character variable of length 10,
(4) factor variable of length 10

```{r}
hw1_df = tibble(
vec_rsamp = runif(n = 10, min = 0, max = 5),
vec_logical = vec_rsamp >= 2,
vec_char = letters[1:10],
vec_factor = rep(c("Male", "Female"), length = 10)
)
hw1_df

```


Now we will try to take the mean for all these variables. 

```{r}
mean(hw1_df[["vec_rsamp"]])
mean(hw1_df[["vec_logical"]])
mean(hw1_df[["vec_char"]])
mean(hw1_df[["vec_factor"]])
```

We are only able to get the mean for vec_rsamp becaue it is a numeric variable. All the others are not numeric variables and so it is not possible to calculate a mean.



