---
title: "Homework 1"
author: "Aakriti Gupta"
date: 2018-09-21
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library (knitr)
library(dplyr)
```

## **Problem 1**
We will first create a dataframe containing variables of four different types: 
(1) random sample of 10 numbers from a uniform distribution between 0 to 5
(2) logical vector indicating whether the numbers from the sample are greater than 2
(3) character variable of length 10,
(4) factor variable of length 10

```{r}
hw1_df = tibble(
vec_rsamp = runif(n = 10, min = 0, max = 5),
vec_logical = vec_rsamp >= 2,
vec_char = letters[1:10],
vec_factor = rep(c("Male", "Female"), length = 10)
)
hw1_df

```

## Calculation of mean
Now we will try to take the mean for all these variables. 

```{r}
mean(hw1_df[["vec_rsamp"]])
mean(hw1_df[["vec_logical"]])
mean(hw1_df[["vec_char"]])
mean(hw1_df[["vec_factor"]])
```

We are only able to get the mean for vec_rsamp becaue it is a numeric variable and vec_logical because it is a logical varoable with value of 0 or 1. All the others are not numeric variables and so it is not possible to calculate a mean.

## Conversion of variables
We will now try to convert the on-numeric variables into numeric variables.

(1) Converting the character variable into numeric variable

```{r}
nvec_char = as.numeric(hw1_df[["vec_char"]])
nvec_char
```

It introduces NAs in place of characters

(2) Converting the logical variable into numeric variable

```{r}
nvec_logical = as.numeric(hw1_df[["vec_logical"]])
nvec_logical
```

It introduces 0 or 1 values based on whether the values were true or false

(3) Converting the factor variable into numeric variable

```{r}
nvec_factor = as.numeric(hw1_df[["vec_factor"]])
nvec_factor
```

It introduces NAs in place of the factorial values


## Second part of variable conversion:

(1) Converting character variable from character to factor to numeric

```{r}
fvec_char = as.factor(hw1_df[["vec_char"]])
fvec_char
nfvec_char = as.numeric(fvec_char)
nfvec_char
```

This code first converts the charactors into factor variables of 10 levels and then finally converts them into numeric values.


(2) Converting factor variable from factor to character to numeric

```{r}
cvec_char = as.character(hw1_df[["vec_factor"]])
cvec_char
ncvec_char = as.numeric(cvec_char)
ncvec_char
```

This code converts the factor variable into character variable which no longer has 0 or 1 values and so it not able to be converted into numeric values.

## **Problem 2**
We will first create a data frame comprised of:
x: a random sample of size 1000 from a standard Normal distribution
y: a random sample of size 1000 from a standard Normal distribution
A logical vector indicating whether the x + y > 0
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector

```{r}
set.seed(1234)
hw1_df2 = tibble(
x = rnorm(1000),
y = rnorm(1000),
z = (x + y) > 0,
nz = as.numeric(z),
fz = as.factor(z)
)
hw1_df2

```

## In-line coding
This dataframe contains a vector with `r length(hw1_df2[["x"]])`  observations. The mean of the variable x is `r mean(hw1_df2[["x"]])` and the median is `r median(hw1_df2[["x"]])`. The proportion of cases for which the logical vector is true is `r mean(hw1_df2[["z"]])`.

## Making the scatterplot

(1) Color points using logical variable

```{r}
ggplot(hw1_df2, aes(x=x, y=y, color=z)) + geom_point()
ggsave("scatter_plot_hw1.pdf", height = 4, width = 6)
```

(2) Color points using numeric variable
```{r}
ggplot(hw1_df2, aes(x=x, y=y, color=nz)) + geom_point()
```

(3) Color points using factor variable
```{r}
ggplot(hw1_df2, aes(x=x, y=y, color=fz)) + geom_point()
```

All the scatter plots essentially look similar because the points colored based only on 2 values - in the case of logical - true vs. false, in the case of numeric - 1 vs. 0 and in the case of factor - "true" vs. "false"


